---
title: "p8105_hw6_sc5154"
author: "Shaohan Chen"
date: "2022-11-30"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
library(modelr)
```


## Problem 1

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

To obtain a distribution for $\hat{r}^2$, we'll do the following steps: 
* Draw bootstrap samples; 
* Generate a model to each
* Extract the value we are concerned with
* Summarize. 
Here, we'll use `modelr::bootstrap` to draw the samples and `broom::glance` to produce `r.squared` values. 

```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + geom_density()
```

In this example, the $\hat{r}^2$ value is high, and the upper bound at 1 may account for the generally skewed shape of the distribution. 

If we want to construct a confidence interval for $R^2$, we could take the 2.5% and 97.5% quantiles of the estimates across bootstrap samples. 

However, because the shape is not symmetric, using the mean +/- 1.96 times the standard error probably wouldn't work well.

We can produce a distribution for $\log(\beta_0 * \beta1)$ using a similar approach, with a bit more wrangling before we make our plot.

```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  rename(beta0 = `(Intercept)`, beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1)) %>% 
  ggplot(aes(x = log_b0b1)) + geom_density()
```

As with $r^2$, this distribution is somewhat skewed and has some outliers. 

The point of this is not to say we should always use the bootstrap -- it's possible to establish "large sample" distributions for strange parameters / values / summaries in a lot of cases, and those are great to have. But it is helpful to know that there's a way to do inference even in tough cases. 


## Problem 2

First we import the original raw dataset.
```{r}
homicide_raw = read_csv("Data/homicide-data.csv")
```

Next, we tidy the dataset according to the requirements. The steps would include:

* Create a `city_state` variable 
* Create a binary variable indicating whether the homicide is solved.
* Omit some cities like Dallas which don't report victim race.
* Omit data entry mistake.
* Limit the analysis those for whom `victim_race` is `white` or `black`.
* Make sure that the `victim_age` is numeric.

In addition, we make sure that `victim_race`, `victim_sex` are factors, for the convenience of further modelling.

```{r}
homicide_df = 
  homicide_raw %>%
  mutate(
    city_state = str_c(city, ", ", state),
    solved = ifelse(disposition %in% c("Closed without arrest", "Open/No arrest"), FALSE, TRUE)
  ) %>%
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))) %>%
  filter(!(city_state %in% c("Tulsa, AL"))) %>%
  filter(victim_race %in% c("White", "Black")) %>%
  mutate(
      victim_age = as.numeric(victim_age),
      victim_sex = as.factor(ifelse(victim_sex == "Unknown", NA, victim_sex)),
      victim_race = as.factor(victim_race)
  )
```

Now the dataset tidy process is completed.

Next, we use `glm` function to fit a logistic regression with resolved vs. unresolved as the outcome and victim age, sex and race as predictors for the city of Baltimore.

Later, we will save the output of the model as an R object, and apply the `broom:tidy` to this object, and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
fit_logistic = 
  homicide_df %>%
  filter(city == "Baltimore") %>%
  glm(solved ~ victim_age + victim_sex + victim_race, data = ., family = binomial())
```

```{r}
fit_logistic %>%
  broom::tidy() %>%
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate + qnorm(0.025) * std.error),
    CI_upper = exp(estimate + qnorm(0.975) * std.error)
  ) %>%
  filter(term == "victim_sexMale") %>%
  select(OR, CI_lower, CI_upper) %>%
  knitr::kable(digits = 4)
```

Now we run `glm` for each of the cities in our dataset, and extract the adjusted odds ratio and CI for solving homicides comparing male victims to female victims. 

We will do this within a “tidy” pipeline, making use of `purrr::map`, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

We first write a function for operating logistic regression for each city. The basic structure is much like what we have written in previous paragraph.
```{r}
glm_training = function(city_data){
  glm_model = glm(solved ~ victim_age + victim_sex + victim_race, data = city_data, family = binomial())
  glm_model %>%
    broom::tidy() %>%
    mutate(
      OR = exp(estimate),
      CI_lower = exp(estimate + qnorm(0.025) * std.error),
      CI_upper = exp(estimate + qnorm(0.975) * std.error)
    ) %>%
    filter(term == "victim_sexMale") %>%
    select(OR, CI_lower, CI_upper)
}
```

```{r}
cities_OR = 
  homicide_df %>%
  select(city_state:solved, victim_race:victim_sex) %>%
  nest(input_data = solved:victim_sex) %>%
  mutate(
    fit_logistic_city = purrr::map(input_data, glm_training)
  ) %>%
  select(city_state, fit_logistic_city) %>%
  unnest(fit_logistic_city)
```

Finally, we make a plot that shows the estimated ORs and CIs for each city. And we will organize cities according to estimated OR, and comment on the plot.

The source data of estimated ORs and CIs look like this:
```{r}
head(cities_OR, 5)
```

The plot is:
```{r}
cities_OR %>%
  mutate(
    city_state = fct_reorder(city_state, OR)
  ) %>%
  ggplot(aes(
    x = city_state,
    y = OR
  )) +
  geom_point() +
  geom_errorbar(aes(
    ymax = CI_upper,
    ymin = CI_lower
  )) +
  labs(
    title = "Odd Ratio's estimate & CI of solving homicides: male victims compared with female victims",
    x = "City, State",
    y = "Odds Ratio"
  ) + 
  theme(
   title = element_text(size = 8)
  )
```




